---
title: "Practical Machine Learning Project - Human Activity Recognition"
output: 
  html_document:
    toc: yes
    keep_md: true
date: '`r format(Sys.time(), "%B %d, %Y")`'
author: by E. Brian Welch using `r sprintf("%s and knitr version %s", 
  R.version.string, packageVersion("knitr"))`
---

```{r setup, echo=FALSE}
require(knitr)
opts_chunk$set(echo = FALSE, cache=TRUE, comment="")
read_chunk('Practical_Machine_Learning_Project.R')
```

```{r Practical_Machine_Learning_Project_setup, echo=FALSE}
```

## Project Overview
The goal of this project is to predict the manner in which a human subject performed a dumbbell (1.25kg) exercise based on accelerometer measurements gathered on a subject's arm, forearm, belt and the dumbbell itself. The Class of excercise included A, B, C, D and E where Class A corresponds to the correct execution of the exercise. The other 4 classes correspond to common mistakes. Exercises were performed by six male participants (age 20-28 years) with limited weight lifting experience. Participants were supervised by an experienced weight lifter to ensure the execution matched the intended simulation manner. Additional study details are available at [http://groupware.les.inf.puc-rio.br/har](http://groupware.les.inf.puc-rio.br/har).

## Data Overview
Data for the project was provided as two comma-separated-variable files `pml-training.csv` (19,622 observations of 160 variables) and `pml-testing.csv` (20 observations of 160 variables). The last column of `pml-training.csv` is the Class (A, B, C, D or E) of the executed dumbbell exercise. The last column of the `pml-testing.csv` is the "problem_id" of the unknown exercise to be predicted by the model created in this project.

## Data Processing and Prediction Results
After loading the orginal testing and training data files, only columns related to belt, arm or dumbbell measurements were kept as predictors. Additionally, any columns containing NA or "#DIV/0!" values were excluded as predictors. Applying these constraints yielded 52 predictors. For cross-validation, the provided training observations were subdivided into 11,778 (60%) train and 7,846 (40%) test observations. Due to its widely known robustness and success in real world applications, a random forest predication model was created.

### 1. Load and Transform Data
```{r Practical_Machine_Learning_Project_section_1, echo=TRUE}
```

### 2. Prepare for Cross-Validation 
```{r Practical_Machine_Learning_Project_section_2, echo=TRUE}
```

### 3. Create a Random Forest Model
```{r Practical_Machine_Learning_Project_section_3, echo=TRUE}
```

### 4. Display Plots Describing the Random Forest Model
```{r Practical_Machine_Learning_Project_section_4, echo=TRUE, fig.width=8, fig.height=8, dpi=120}
```

### 5. Estimate Out of Sample Error Using Cross-Validation
```{r Practical_Machine_Learning_Project_section_5, echo=TRUE}
```

### 6. Predict Using pml-test Data
```{r Practical_Machine_Learning_Project_section_6, echo=TRUE}
```

### 7. Write pml-test Predictions to Text Files
```{r Practical_Machine_Learning_Project_section_7, echo=TRUE}
```

## Summary
The created random forest model used 500 trees and achieved classification error of approximately 1% or better for all 5 classes. The most important predictors in the random forest model were `yaw_belt` and `roll_belt`. The estimated OOB (out-of-bag) error was 0.66% which agreed well with the separately cross-validated out of sample error rate estimation of 0.59%. The predictions for the 20 cases in the `pml-testing.csv` file scored 100% correct.